---
title: "metafor-resumido"
author: "Clarissa F. D. Carneiro"
date: "2022-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(metafor)
library(metaviz)
library(glmulti)
```

```{r}
dados = read_xlsx("dados_todos-resumido.xlsx")
```

Para análises, precisamos da média e n dos dois grupos e da variacao no tratado, seja SEM, SD ou Unclear. Aqui excluo os dados incompletos. 
```{r}
dados_completos = dados %>% filter(control_mean!=0&!is.na(control_mean)&treated_mean!=0&!is.na(treated_mean)&control_n!=0&!is.na(control_n)&treated_n!=0&!is.na(treated_n)&treated_variation!=0&!is.na(treated_variation))
```

Calculei SDs a partir de SEMs (se estava 'unclear', considerei SEM).
```{r}
dados_completos = dados_completos %>% mutate(control_sd = 
                                       if_else(condition = variation=="SD", true = control_variation, false = (control_variation*sqrt(control_n))))

dados_completos = dados_completos %>% mutate(treated_sd = 
                                       if_else(condition = variation=="SD", true = treated_variation, false = (treated_variation*sqrt(treated_n))))

```

Como temos comparacoes dois tipos diferentes de normalizacao (com e sem variacao no controle), preciso calcular o tamanho de efeito separadamente.

```{r}
dados_two_sample = dados_completos %>% filter(!is.na(control_sd)) %>% filter(control_sd!=0)

dados_one_sample = dados_completos %>% filter(is.na(control_sd)|control_sd==0)
```

```{r}
#usando a variacao de cada grupo
dados_two_sample_smd = escalc(measure = "SMD", m1i = as.numeric(treated_mean), m2i = as.numeric(control_mean), sd1i = as.numeric(treated_sd), sd2i = as.numeric(control_sd), n1i = as.numeric(treated_n), n2i = as.numeric(control_n), data = dados_two_sample)

#variacao do tratado em ambos os grupos
dados_one_sample_smd = escalc(measure = "SMD", m1i = as.numeric(treated_mean), m2i = as.numeric(control_mean), sd1i = as.numeric(treated_sd), sd2i = as.numeric(treated_sd), n1i = as.numeric(treated_n), n2i = as.numeric(control_n), data = dados_one_sample)

#combinando os dados de volta
dados_meta_smd = rbind(dados_two_sample_smd, dados_one_sample_smd)
```


Meta-análise (2-level SMD)
```{r}
meta1 = rma(yi=yi, vi=vi, data = dados_meta_smd, measure = "SMD", method = "REML", slab = rayyan.key)
summary(meta1)
confint(meta1)
```

Forest plot (SMD)
```{r, fig.height=15, fig.width=10}
forest1 = viz_forest(meta1, study_labels = dados_meta_smd$rayyan.key, col = "black", annotate_CI = TRUE, xlab = "Hedge's g")
forest1
```

Meta-análise (3-level SMD)
```{r}
dados_meta_smd = dados_meta_smd %>% mutate(Comparison.ID = 1:nrow(dados_meta_smd))

meta_3l = rma.mv(yi=yi, V=vi, data = dados_meta_smd, method = "REML", random = ~1|rayyan.key/Comparison.ID)
meta_3l
confint(meta_3l)
```

```{r}
#As there is not function in the package metafor for estimating I^2 values for 3-level models, we imported the function from Mathias Harrer & David Daniel Ebert (https://raw.githubusercontent.com/MathiasHarrer/dmetar/master/R/mlm.variance.distribution.R)

mlm.variance.distribution = var.comp = function(x){

  m = x

  # Check class
  if (!(class(m)[1] %in% c("rma.mv", "rma"))){
    stop("x must be of class 'rma.mv'.")
  }

  # Check for three level model
  if (m$sigma2s != 2){
    stop("The model you provided does not seem to be a three-level model. This function can only be used for three-level models.")
  }

  # Check for right specification (nested model)
  if (sum(grepl("/", as.character(m$random[[1]]))) < 1){
    stop("Model must contain nested random effects. Did you use the '~ 1 | cluster/effect-within-cluster' notation in 'random'? See ?metafor::rma.mv for more details.")
  }

  # Get variance diagonal and calculate total variance
  n = m$k.eff
  vector.inv.var = 1/(diag(m$V))
  sum.inv.var = sum(vector.inv.var)
  sum.sq.inv.var = (sum.inv.var)^2
  vector.inv.var.sq = 1/(diag(m$V)^2)
  sum.inv.var.sq = sum(vector.inv.var.sq)
  num = (n-1)*sum.inv.var
  den = sum.sq.inv.var - sum.inv.var.sq
  est.samp.var = num/den

  # Calculate variance proportions
  level1=((est.samp.var)/(m$sigma2[1]+m$sigma2[2]+est.samp.var)*100)
  level2=((m$sigma2[2])/(m$sigma2[1]+m$sigma2[2]+est.samp.var)*100)
  level3=((m$sigma2[1])/(m$sigma2[1]+m$sigma2[2]+est.samp.var)*100)

  # Prepare df for return
  Level=c("Level 1", "Level 2 (exp)", "Level 3 (art)")
  Variance=c(level1, level2, level3)
  df.res=data.frame(Variance)
  colnames(df.res) = c("% of total variance")
  rownames(df.res) = Level
  I2 = c("---", round(Variance[2:3], 2))
  df.res = as.data.frame(cbind(df.res, I2))

  totalI2 = Variance[2] + Variance[3]


  # Generate plot
  df1 = data.frame("Level" = c("Sampling Error", "Total Heterogeneity"),
                  "Variance" = c(df.res[1,1], df.res[2,1]+df.res[3,1]),
                  "Type" = rep(1,2))

  df2 = data.frame("Level" = rownames(df.res),
                   "Variance" = df.res[,1],
                   "Type" = rep(2,3))

  df = as.data.frame(rbind(df1, df2))


  g = ggplot(df, aes(fill=Level, y=Variance, x=as.factor(Type))) +
    coord_cartesian(ylim = c(0,1), clip = "off") +
    geom_bar(stat="identity", position="fill", width = 1, color="black") +
    scale_y_continuous(labels = scales::percent)+
    theme(axis.title.x=element_blank(),
          axis.text.y = element_text(color="black"),
          axis.line.y = element_blank(),
          axis.title.y=element_blank(),
          axis.line.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.y = element_line(lineend = "round"),
          legend.position = "none",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          legend.background = element_rect(linetype="solid",
                                           colour ="black"),
          legend.title = element_blank(),
          legend.key.size = unit(0.75,"cm"),
          axis.ticks.length=unit(.25, "cm"),
          plot.margin = unit(c(1,3,1,1), "lines")) +
    scale_fill_manual(values = c("darkseagreen3", "deepskyblue3", "darkseagreen2",
                                 "deepskyblue1", "deepskyblue2")) +

    # Add Annotation

    # Total Variance
    annotate("text", x = 1.5, y = 1.05,
             label = paste("Total Variance:",
                           round(m$sigma2[1]+m$sigma2[2]+est.samp.var, 3))) +

    # Sampling Error
    annotate("text", x = 1, y = (df[1,2]/2+df[2,2])/100,
             label = paste("Sampling Error Variance: \n", round(est.samp.var, 3)), size = 3) +

    # Total I2
    annotate("text", x = 1, y = ((df[2,2])/100)/2-0.02,
             label = bquote("Total"~italic(I)^2*":"~.(round(df[2,2],2))*"%"), size = 3) +
    annotate("text", x = 1, y = ((df[2,2])/100)/2+0.05,
             label = paste("Variance not attributable \n to sampling error: \n", round(m$sigma2[1]+m$sigma2[2],3)), size = 3) +

    # Level 1
    annotate("text", x = 2, y = (df[1,2]/2+df[2,2])/100, label = paste("Level 1: \n",
                                                                       round(df$Variance[3],2), "%", sep=""), size = 3) +

    # Level 2
    annotate("text", x = 2, y = (df[5,2]+(df[4,2]/2))/100,
             label = bquote(italic(I)[Level2]^2*":"~.(round(df[4,2],2))*"%"), size = 3) +

    # Level 3
    annotate("text", x = 2, y = (df[5,2]/2)/100,
             label = bquote(italic(I)[Level3]^2*":"~.(round(df[5,2],2))*"%"), size = 3)

  returnlist = list(results = df.res,
                    totalI2 = totalI2,
                    plot = g)
  class(returnlist) = c("mlm.variance.distribution", "list")

  invisible(returnlist)

  returnlist

}
```

```{r fig.keep='none', warning=FALSE}
mlm.variance.distribution(x = meta_3l)
```

Meta-regressions (3-level)

Differentiation
```{r}
meta_differentiation = rma.mv(yi=yi, V=vi, data = dados_meta_smd, method = "REML", random = ~1|rayyan.key/Comparison.ID, mods = ~relevel(factor(dados_meta_smd$Diferentiation_method), ref="No differentiation"))

meta_differentiation
```

Differentiation bubble plot
```{r}
ggplot(dados_meta_smd, aes(x = Diferentiation_method, y=yi, size = 1/sqrt(vi))) + 
  geom_point(shape = 1, position =  position_jitterdodge(dodge.width = 0), stroke = 1) + 
  labs(y = "Hedge's g", x = "Differentiation method") + 
  theme_classic() + 
  scale_size_continuous(guide = "none") + 
  theme(legend.position = "bottom") + 
  scale_y_continuous(n.breaks = 7) + 
  geom_abline(yintercept = 0, slope = 0, linetype = "dashed", color = "grey")
```

Diff. duration
```{r}
meta_difduration = rma.mv(yi=yi, V=vi, data = dados_meta_smd, method = "REML", random = ~1|rayyan.key/Comparison.ID, mods = ~dados_meta_smd$Diferentiation_duration_days)

meta_difduration
```

Diff. duration bubble plot
```{r}
ggplot(dados_meta_smd, aes(x = Diferentiation_duration_days, y=yi, size = 1/sqrt(vi))) + 
  geom_point(shape = 1, stroke = 1) + 
  labs(y = "Hedge's g", x = "Duration of differentiation (days)") + 
  theme_classic() + 
  scale_size_continuous(guide = "none") + 
  theme(legend.position = "bottom") + 
  scale_y_continuous(n.breaks = 7) + 
  scale_x_continuous(n.breaks = 7) + 
  geom_abline(slope = 0, linetype = "dashed", color = "grey")
```
